The \dealii{} Library, Version 9.3

### 2021-06-18 9.3/paper.tex
Edit the candi section.

The major changes of this release are:
%
\begin{itemize}
  \item Experimental support for simplex and mixed meshes (see Section~\ref{subsec:simplex});
  \item Improved flexibility of the particle infrastructure (see Section~\ref{subsec:particles});
  \item Support for global-coarsening multigrid algorithms (see Section~\ref{subsec:mg});
  \item Advances in the matrix-free infrastructure  (see Section~\ref{subsec:mf});
  \item Usage of MPI-3.0 shared-memory features to reduce memory footprint  (see Section~\ref{subsec:sm});
  \item Improved support for evaluation and integration at arbitrary points (see Section~\ref{subsec:fepointvalues});
  \item Nine new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
\end{itemize}
%
In addition, we discuss the \texttt{candi} installation program in Section~\ref{subsec:candi}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Source based toolchain installer}
\label{subsec:candi}

The requirement to download, compile and install \dealii{} and its dependencies from
source is of major interest for many \dealii{}-users.
%
Compiling dependencies from source can be of importance on arbitrary Linux-based operating systems for which
binary packages aren't available or on compute clusters and other machines
without root privileges for the user to install system dependencies.

The source based installation of \dealii{} and its many dependent libraries
can be done with the \candi{} script tool for various
Linux operating systems, within the Windows Subsystem Linux (WSL) and on OSX (experimental).
%
The general assumption is that a C, C++ and Fortran compiler and suitable MPI-compilers
for the base compilers as well as the corresponding development system packages are
available.

\candi is a bash-script based tool, it is the abbreviation of ``compile and install''
and it is released under the GNU LGPL v3.0.
%
The origin of \candi{} is a fork made in 2013 from \texttt{dorsal},
a retired source based installer for the \texttt{FEniCS} library.
%
The \candi{} tool can be found via the download page of \dealii{} since version 8.5
or from github.com/dealii/candi and is under active development.
To install older releases of \dealii{} toolchains one can checkout the branch
corresponding to the \dealii{}-version with \texttt{git}.

The installer supports the user to download, unpack, compile and install a certain
library, a list of libraries or a complete toolchain.
A library is called a package within \candi{}.
The toolchain installation is the default behaviour and the default configuration
ensures that most of the \dealii{} step tutorials can be used directly.
The package installation mode can be handy to generate docker containers
for instance.
Especially in the toolchain mode we can check for dependencies from one library to
another and handle them accordingly.

Each package for a library is defined by variables for its name and version,
a (remote or local) download location,
its packaging format (e.g. tar, gz or zip),
a checksum to prevent from incomplete downloads,
its general build chain (e.g. using \texttt{autotools}, \texttt{cmake} and others)
as well as configuration options.
%
Moreover, one can give specific instructions for each of the steps being
necessary for the unpacking, configuration or building of the package.

An internet connection isn't necessary
if the source files are downloaded and
repositories are checked out to its unpacking location,
e.g. from a previous run on a machine having an internet connection.
%
The download of a package can be supported via local or remote mirror servers
similar as we are doing it already for several of the packages.
%
If the temporary files of \candi{} aren't removed, one can use the developer
mode to prepare patches or developments for any of the packages.
%
A new feature allows \candi{} to skip user prompts which can be used for
a batch mode run on a compute cluster job for instance.

This source based installer is implemented in a modular fashion.
The \texttt{candi.sh} script is implemented without any knowledge
on a package or toolchain for a specific operating system.
Therefore, a couple of checks and setting internal variables
for known platforms are done.

The \candi{} source based installer is organised in the following way:
\begin{itemize}
\item \texttt{candi.sh}: the bash-script which controls the overall process.
The command line options are given by the output of ``\texttt{candi.sh -h}''
and explained in the \texttt{README.md} file.

\item \texttt{candi.cfg} (or \texttt{local.cfg}): the default toolchain
configuration file. Here one can easily switch on/off important features,
give additional configuration options to \dealii{} and
define the set of packages for the toolchain.

\item \texttt{deal.II-toolchain} folder: the central project folder for
\dealii{}-based toolchains.
It contains subfolders
for packages (libraries, dependencies, general tools in a specific version),
platforms (operating systems) and
for patches to be applied after unpacking a package.

\end{itemize}

The \texttt{dealii/candi} \texttt{github} repository includes continuous
integration (CI) to ensure that new features of the script or new packages
are working as expected.

>>>>

The major changes of this release are:
%
\begin{itemize}
  \item Experimental support for simplex and mixed meshes (see Section~\ref{subsec:simplex});
  \item Improved flexibility of the particle infrastructure (see Section~\ref{subsec:particles});
  \item Support for global-coarsening multigrid algorithms (see Section~\ref{subsec:mg});
  \item Advances in the matrix-free infrastructure  (see Section~\ref{subsec:mf});
  \item Usage of MPI-3.0 shared-memory features to reduce memory footprint  (see Section~\ref{subsec:sm});
  \item Improved support for evaluation and integration at arbitrary points (see Section~\ref{subsec:fepointvalues});
  \item Nine new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
\end{itemize}
%
In addition, we discuss the \candi{} installation program in Section~\ref{subsec:candi}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The ource based toolchain installer \candi{}}
\label{subsec:candi}

The requirement to download, compile, and install \dealii{} and its dependencies from
source is a major obstacle to many \dealii{} users.
Compiling all dependencies from source can be difficult on operating systems for which
binary packages aren't available or on compute clusters and other machines
without root privileges for the user to install system dependencies.

The source based installation of \dealii{} and its many dependent libraries
can be done with the \candi{} script tool for various
Linux operating systems, within the Windows Subsystem Linux (WSL) and
on OS X (experimental).
%
The general assumption is that a C, C++ and Fortran compiler and suitable MPI-compilers
for the base compilers as well as the corresponding development system packages are
available.

\candi is a bash-script based tool, is an abbreviation of ``compile and install''
and is released under the GNU LGPL v3.0.
%
The origin of \candi{} is a fork made in 2013 from \texttt{dorsal},
a now-retired source based installer for the \texttt{FEniCS} library.
%
The \candi{} tool can be found via the download page of \dealii{} since version 8.5
or from \url{github.com/dealii/candi}, and is under active development.
To install older releases of \dealii{} toolchains, one can check out
the \candi{} branch of the \texttt{git} repository that
corresponds to the desired \dealii{} version.

\candi{} downloads, unpacks, compiles, and installs an individual
library (called a ``package'' in \candi{}), a list of libraries, or a complete toolchain.
The toolchain installation is the default behavior and the default configuration
ensures that most of the \dealii{} step tutorials can be used directly.
The package installation mode is also useful to generate docker containers.
In toolchain mode, \candi{} checks for and deals appropriately with
dependencies between libraries.

Each package for a library is defined by variables for its name and version,
a (remote or local) download location,
its packaging format (e.g. tar, gz, or zip),
a checksum,
its general build chain (e.g., using \texttt{autotools}, \texttt{cmake}, or others),
as well as configuration options.
Moreover, one can give specific instructions for each of the steps being
necessary for the unpacking, configuration, or building of a package.
A new feature allows \candi{} to skip user prompts, allowing its use in
a batch mode.

\candi{} does all this by either downloading packages from the
internet, including through mirrors, or re-using previously downloaded
or checked out from a repository.
If temporary files of \candi{} are not removed, one can use the developer
mode to prepare patches or developments for any of the packages.

\candi{} is organized in the following way:
\begin{itemize}
\item \texttt{candi.sh}: the bash script which controls the overall process.
Available command line options are listed by calling ``\texttt{candi.sh -h}''
and explained in the \texttt{README.md} file.

\item \texttt{candi.cfg} (or \texttt{local.cfg}): the default toolchain
configuration file. Here one can easily switch on/off features,
give additional configuration, and
list the packages for a toolchain.

\item \texttt{deal.II-toolchain} folder: the central project folder for
\dealii{}-based toolchains.
It contains subfolders
for packages (libraries, dependencies, general tools in a specific version),
platforms (operating systems) and
for patches to be applied after unpacking a package.
\end{itemize}

The installer is developed in the \texttt{dealii/candi}
\texttt{github} repository, and uses continuous
integration (CI) to ensure that old and new features
are working as expected.
